Dashboard section

### 2022-07-29 Report/Draft Dashboard section.Rmd
Just cleaning.
Updating. Adding file reading code.

>Ca. 8 sider.

@@

If the validation is successful the files are read using the functions developed for this purpose.

@@

The top row contains two plots showing the annual result and balance data for the years covered by Proff.no. These are fairly simple bar plots, including a `fill` parameter colouring the bars depending on whether the numbers are positive or negative. They do, however, rely on two simple additional functions, one which checks whether the currency code is uniform across the years, and, if so, using it for the y scale, or, otherwise, uses a question mark instead of a currency code, both using `scales::label_dollar()`.
```{r dashboard_currency, eval=FALSE}
# Reactive function to check for currency and use currency code or ?
currency_scale <- reactive({
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(htmldata()$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(htmldata()$valutakode[1], " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )
})
```

@@

# Reactive function to create plotly caption (ggplot captions and subtitles
# are lost in the conversion) (shamelessly stolen from the Internet, with
# minor modifications).
# Checks for uniform currency code and prints a warning
# if there are multiple currency codes.
plotly_caption <- reactive({
  list(
    x = 1, y = -0.1, text = ifelse(length(unique(htmldata()$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency"
    ),
    showarrow = F, xref = "paper", yref = "paper",
    xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
    font = list(size = 11)
  )
})
```

@@

The revenue and expenses plots rely on a separate 'maximum' function to create the y scales; this is used so that both are on the same scale, and to derive the upper limit of the scale from the supplied data. The lower limit is derived from the supplied data in each plot in order not to suppress any erroneous totals. (This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function reads the Excel sheet anew in order to make use of the Category column:
```{r dashboard_maximum, eval=FALSE}
# Reactive function to create the maximum value for revenue and expenses plots
maximum <- reactive({
  # reading the Excel sheet anew
  read_excel(input$excel$datapath) |>
    # removing the firm and sub-category columns
    select(-Firm, -`Sub-category`) |>
    # filtering to retain only revenue and expenses entries
    filter(Category == "Revenue" | Category == "Expenses") |>
    # grouping and summarising to calculate maximum value
    group_by(Year, Category) |>
    summarise(
      sum = sum(Value, na.rm = TRUE)
    ) |>
    # Creating list of sums
    pull() |>
    # selecting the highest value
    max()
})
```

@@

Finally, the dashboard contains a separate tab with some basic details about the KPIs and plots. At the top it includes a heading containing the name of the NGO as written in the Excel file:
```{r dashboard_title_creation, eval=FALSE}
# NGO title ==================================================================
NGO <- reactive({
  # validating the file again
  validate(need(
    tools::file_ext(input$excel$datapath) == "xlsx",
    "Please upload an Excel file."
  ))
  # fetching the name
  read_excel(input$excel$datapath) |>
    # Turning the needed column into a list
    pull(Firm) |>
    # discarding all NAs in case the name is somewhere
    purrr::discard(is.na) |>
    # choosing the first available name
    first()
})

>>>>

>Ca. 11 sider med all koden...

@@

If the validation is successful the files are read using the functions developed for this purpose, which conduct further testing, outputting error messages using `validate()`:
```{r dashboard_file_functions, eval=FALSE}
# Data reading functions =======================================================

## Functions to read html file -------------------------------------------------
html_test_create <- function(file, start, end) {
  # Reading file and node
  htmltest <- read_html(file, encoding = "UTF-8") |>
    html_node("#inner-frame")

  # Checking if node exists
  if (class(htmltest) == "xml_node") {
    htmltable <- htmltest |>
      html_table()
  } else {
    validate("Please upload the specified type of html document.")
  }

  # Basic check if the table is correct
  if ("REGNSKAPSPERIODE" %in% names(htmltable)) {
    htmlready <- htmltable
  } else {
    validate("Please upload the specified type of html document.")
  }

  # Testing if the correct years have been entered
  if (as.character(all_of(start)) %in% names(htmlready) &&
    as.character(all_of(end)) %in% names(htmlready)) {
    # Creating the data frame
    read_proff_one(htmltable, start, end)
  } else {
    validate("Please ensure you have entered the correct start and end years.")
  }
}

read_proff_one <- function(object, start, end) {
  object |>
    # removing empty column, using all_of(argument)
    select(
      REGNSKAPSPERIODE:all_of(start)
    ) |>
    # removing white space
    mutate(
      REGNSKAPSPERIODE = str_squish(REGNSKAPSPERIODE)
    ) |>
    # removing duplicate table sometimes created by read_html()
    filter(
      !row_number() > 176
    ) |>
    # removing duplicate rows
    filter(
      !grepl("Lukk", REGNSKAPSPERIODE)
    ) |>
    # tidying data using pivot_longer() and all_of(argument)
    pivot_longer(
      all_of(end):all_of(start),
      names_to = "year"
    ) |>
    # creating a currency code column and turning current currency codes into
    # NAs
    mutate(
      # changing to real NAs
      value = na_if(value, "-"),
      # turning years into numbers
      year = as.integer(year),
      # creating a currency code column and turning current currency codes into
      # NAs
      valutakode = ifelse(REGNSKAPSPERIODE == "Valutakode", value, NA)
    ) |>
    fill(
      valutakode,
      .direction = "downup"
    )
}

read_proff_two <- function(object) {
  object |>
    # removing dates and redundant currency
    select(-valutakode) |>
    filter(
      REGNSKAPSPERIODE != "Sluttdato" & REGNSKAPSPERIODE != "Startdato" &
        REGNSKAPSPERIODE != "Valutakode"
    ) |>
    mutate(
      # removing white space in numbers
      value = str_replace_all(value, "\\s", ""),
      # turning numbers into true numbers
      value = as.numeric(value),
      # removing values that are years
      value = ifelse(grepl("i hele 1000", REGNSKAPSPERIODE), NA, value)
    ) |>
    # removing any remaining duplicate rows
    distinct() |>
    # pivoting to create 'tidy' variables
    pivot_wider(
      names_from = REGNSKAPSPERIODE, values_from = value
    ) |>
    # tidying names
    clean_names() |>
    # removing headings from the table
    select(
      -lederlonn_i_hele_1000,
      -resultatregnskap_i_hele_1000,
      -balanseregnskap_i_hele_1000
    ) |>
    # making it clear what lonn refers to
    rename(
      lederlonn = lonn
    ) |>
    # Creating new variables for plots
    mutate(
      annual_profit = ifelse(arsresultat >= 0, TRUE, FALSE),
      balance_positive = ifelse(sum_egenkapital_og_gjeld >= 0, TRUE, FALSE)
    ) |>
    # arranging by year for practical display
    arrange(desc(year))
}

## Functions to read Excel file ------------------------------------------------
widen_excel_test <- function(file) {
  # Reading Excel file
  read_excel(file) |>
    # cleaning names to enable later operations
    clean_names()
}

# Function to create primary Excel data frame
widen_excel <- function(object) {
  # testing if all columns are present
  if (setequal(
    names(object),
    c("year", "firm", "category", "sub_category", "name", "value")
  )) {
    object |>
      # uniting category and name to deal with multiple 'Other'
      unite(
        entry, c(category, name)
      ) |>
      # deleting Sub-category and Firm
      select(-sub_category, -firm) |>
      # pivoting to create 'tidy' variables
      pivot_wider(names_from = entry, values_from = value) |>
      # cleaning names
      clean_names()
    # Notification if all columns are not present
  } else {
    validate("Please upload data using the specified Excel template.")
  }
}
```

@@

The top row contains two plots showing the annual result and balance data for the years covered by Proff.no. These are fairly simple bar plots, including a `fill` parameter colouring the bars depending on whether the numbers are positive or negative. They do, however, rely on two simple additional functions, one which checks whether the currency code is uniform across the years, and, if so, using it for the y scale, or, otherwise, uses a question mark instead of a currency code, both using `scales::label_dollar()`.
```{r dashboard_currency, eval=FALSE}
# Reactive function to check for currency and use currency code or ?
currency_scale <- reactive({
  scale_y_continuous(
    # Checking for uniform currency code
    labels = ifelse(length(unique(html_currency()$valutakode)) == 1,
      # Using the currency code if uniform
      label_dollar(prefix = str_c(unique(html_currency()$valutakode), " ")),
      # alternative prefix if multiple currencies in the data
      label_dollar(prefix = "? ")
    )
  )
})
```

@@

# Reactive function to create plotly caption (ggplot captions and subtitles
# are lost in the conversion) (shamelessly stolen from the Internet, with
# minor modifications).
# Checks for uniform currency code and prints a warning
# if there are multiple currency codes.
plotly_caption <- reactive({
  list(
    x = 1, y = -0.1,
    text = ifelse(length(unique(html_currency()$valutakode)) == 1,
      "", "Note: not all values are denominated in the same currency"
    ),
    showarrow = F, xref = "paper", yref = "paper",
    xanchor = "right", yanchor = "auto", xshift = 0, yshift = 0,
    font = list(size = 11)
  )
})
```

@@

The revenue and expenses plots rely on a separate 'maximum' function to create the y scales; this is used so that both are on the same scale, and to derive the upper limit of the scale from the supplied data. The lower limit is derived from the supplied data in each plot in order not to suppress any erroneous totals. (This is not a complete data integrity test, but should convey a slight warning if the expenses have been entered as negative numbers.) This function reads the Excel sheet anew in order to make use of the Category column:
```{r dashboard_maximum, eval=FALSE}
# Reactive function to create the maximum value for revenue and expenses plots
maximum <- reactive({
  if (datatest() == TRUE) {
    NA
  } else {
    # reading the Excel sheet anew
    read_excel(input$excel$datapath) |>
      # removing the firm and sub-category columns
      select(-Firm, -`Sub-category`) |>
      # filtering to retain only revenue and expenses entries
      filter(Category == "Revenue" | Category == "Expenses") |>
      # grouping and summarising to calculate maximum value
      group_by(Year, Category) |>
      summarise(
        sum = sum(Value, na.rm = TRUE)
      ) |>
      # Creating list of sums
      pull() |>
      # selecting the highest value
      max()
  }
})
```

@@

Finally, the dashboard contains a separate tab with some basic details about the KPIs and plots. At the top it includes a heading containing the name of the NGO as written in the Excel file:
```{r dashboard_title_creation, eval=FALSE}
# NGO title ==================================================================
NGO <- reactive({
  # checking for the required column again
  if ("firm" %in% names(excel_test())) {
    # fetching the name
    excel_test() |>
      # Turning the needed column into a list
      pull(firm) |>
      # discarding all NAs in case the name is somewhere
      purrr::discard(is.na) |>
      # choosing the first available name
      first()
  } else {
    validate("Please upload data using the specified Excel template.")
  }
})
